---
title: "SNP, SSRs Abies IUFRO 2019"
author: "Nikos Tourvas"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
bibliography: library.bib
output:
  html_notebook:
    code_folding: hide
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  html_document:
    code_folding: hide
    theme: united
    toc: yes
    ## toc_float: yes
  pdf_document:
    toc: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE, echo=FALSE, fig.path = "/figs/")
```

## To Implement

- Information capacity
  - locus numbers
  - polymorphism level
  - gene action mode ???
- Comparison with Krajmerova (only 2 common loci - no custom anal needed)
- Comparison between cohorts (gen div higher in mature???)
  - na, ne, LGP, I 
```{r}
library(popprxl)
library(tidyverse)
```

## Load SNP data set
```{r}
source("import_snp_abies.R")
```

```{r}
threshold_loci <- 0.99
threshold_ind <- 0.99
#obj_SNP <- import.snp.abies2(threshold_ind, threshold_loci, maf=0)
maf <- 0
obj_SNP <- import.snp.abies(threshold_ind, threshold_loci, maf)
```

<!-- ## subset greek pops -->
```{r}
obj_SNP <- popsub(obj_SNP, sublist = c("GR_A", "GR_NR"))
```

<!-- ## Filter specifically for greek pops -->
```{r}
obj_SNP <- missingno(obj_SNP, type="loci",
                     cutoff = 0.1)
obj_SNP <- missingno(obj_SNP, type = "geno",
                     cutoff = 0.1)
obj_SNP <- informloci(obj_SNP, MAF = 0)

```


### Data set details after data filtering:
```{r include=T}
paste(nInd(obj_SNP), "individuals", "and", nLoc(obj_SNP), "loci", sep=" ")
```

## Load SSR data set
```{r}
library(popprxl)

obj_SSR <- read.genalexcel(
  "../data/LGM_DE_SI_GR_final.xlsx",   # name of excel file
  sheet = "Abies_IUFRO2019",             # name of sheet where the genotypes reside
  genclone = F) 


splitStrata(obj_SSR) <- ~Country/Pop
```

<!-- ##subset SSR dataset -->
```{r}
obj_SSR <- popsub(obj_SSR,
                  sublist = c("GR_A", "GR_NR"))
```

## SSRs Private alleles per cohort - level Country/Pop
```{r private_SSR, cache=TRUE, dpi=300, fig.width=8.5, include=T}
private <- poppr::private_alleles(obj_SSR, report = "data.frame")

if(is.data.frame(private)){
  
ggplot(private) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per population using SSR markers") +
  scale_fill_viridis_c()
  
}
```

```{r}
PA_SSR <- as.data.frame(
          rowSums(poppr::private_alleles(obj_SSR, count.alleles=F))
)

PA_SSR$marker <- c("11 SSR loci")

PA_SSR <- PA_SSR %>% 
  rename("Private alleles" = `rowSums(poppr::private_alleles(obj_SSR, count.alleles = F))`)

PA_SSR$Cohort <- rownames(PA_SSR)
```

## SNPs Private alleles per cohort - level Country/Pop
```{r private_SNP, cache=TRUE, dpi=300, fig.width=8.5, include=T}
private <- poppr::private_alleles(obj_SNP, report = "data.frame")

if(is.data.frame(private)){
  
ggplot(private) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per population using SSR markers") +
  scale_fill_viridis_c()
  
}
```

```{r}
PA_SNP <- as.data.frame(
          rowSums(poppr::private_alleles(obj_SNP, count.alleles=F))
)

PA_SNP$marker <- c("140 SNP loci")

PA_SNP <- PA_SNP %>% 
  rename("Private alleles" = `rowSums(poppr::private_alleles(obj_SNP, count.alleles = F))`)

PA_SNP$Cohort <- rownames(PA_SNP)
```

## PA
```{r}
#bind dfs
PA_df <- rbind(PA_SSR, PA_SNP)

pa <- PA_df %>% 
      ggplot(aes(x=Cohort, y=`Private alleles`, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_wrap(~ marker) +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) +
      scale_y_continuous(breaks = c(0, 3, 6, 9, 11))

pa + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11))

# SHOW WHAT IS THE SAMPLE SIZE!!!
```

## PCA SSR

```{r}
library(factoextra)
library(ggord)
```

```{r pca_ssr, cache=T, fig.width=5, fig.height=3, include=T}
## In case there are missing data, replace with mean
X <- tab(obj_SSR, freq = TRUE, NA.method = "mean") 

pca.obj_SSR <- dudi.pca(X, scale = FALSE, scannf = FALSE, nf = 3)

fviz_eig(pca.obj_SSR, main = "Scree plot of PCA")
```

```{r pca_ssr_plot, cache=T, include=T}
ggord(pca.obj_SSR, pop(obj_SSR), arrow = NULL, txt = NULL, 
      grp_title="Populations", ellipse=F, poly=FALSE, size=2, 
      alpha=.8, coord_fix=F, axes = c("1", "2")) +
  geom_hline(yintercept=0,linetype=2) + 
  geom_vline(xintercept=0,linetype=2) +
  ggtitle("Principal Component Analysis") +
  scale_shape_manual('Groups', values = 0:5) +
  scale_color_manual('Groups', values = 
                       c('#1b9e77','#d95f02','#7570b3',
                         '#e7298a','#66a61e','#e6ab02') )

# ggord(pca.obj_SSR, pop(obj_SSR), arrow = NULL, txt = NULL, 
# grp_title="Populations", ellipse=T, poly=FALSE, size=2, 
# alpha=.8, coord_fix=F, axes = c("1", "3")) +
#   geom_hline(yintercept=0,linetype=2) + 
#   geom_vline(xintercept=0,linetype=2) +
#   ggtitle("Principal Component Analysis") +
#   scale_shape_manual('Groups', values = 0:5) +
#   scale_color_manual('Groups', values = 
#                        c('#1b9e77','#d95f02','#7570b3',
#                          '#e7298a','#66a61e','#e6ab02') )
```

## PCA SNP
```{r pca_snp, cache=T, fig.width=5, fig.height=3, include=T}
## In case there are missing data, replace with mean
X <- tab(obj_SNP, freq = TRUE, NA.method = "mean") 

pca.obj_SNP <- dudi.pca(X, scale = FALSE, scannf = FALSE, nf = 3)

fviz_eig(pca.obj_SNP, main = "Scree plot of PCA")
```

```{r pca_snp_plot, cache=T, include=T}
ggord(pca.obj_SNP, pop(obj_SNP), arrow = NULL, txt = NULL, 
      grp_title="Populations", ellipse=F, poly=FALSE, size=2, 
      alpha=.8, coord_fix=F, axes = c("1", "2")) +
  geom_hline(yintercept=0,linetype=2) + 
  geom_vline(xintercept=0,linetype=2) +
  ggtitle("Principal Component Analysis") +
  scale_shape_manual('Groups', values = 0:5) +
  scale_color_manual('Groups', values = 
                       c('#1b9e77','#d95f02','#7570b3',
                         '#e7298a','#66a61e','#e6ab02') )

# ggord(pca.obj_SNP, pop(obj_SNP), arrow = NULL, txt = NULL, 
# grp_title="Populations", ellipse=T, poly=FALSE, size=2, 
# alpha=.8, coord_fix=F, axes = c("1", "3")) +
#   geom_hline(yintercept=0,linetype=2) + 
#   geom_vline(xintercept=0,linetype=2) +
#   ggtitle("Principal Component Analysis") +
#   scale_shape_manual('Groups', values = 0:5) +
#   scale_color_manual('Groups', values = 
#                        c('#1b9e77','#d95f02','#7570b3',
#                          '#e7298a','#66a61e','#e6ab02') )
```

## Summary statistics
```{r summary_statistics, cache=T}

## functions

###### basic statistics and their SEs
table_out <- function(obj, variable, name){
  
means <- colMeans(variable, na.rm=T)
out <- c(means, mean(variable, na.rm = T))

out <- as.data.frame(out)
Pops <- c(popNames(obj), "Total")
rownames(out) <- Pops
colnames(out) <- name

sem_out <- apply(variable, 2, function(x) sd(x) / sqrt(length(x))) 
  ## 2 means work along columns
sem_out_mean <- sd(variable) / sqrt(length(variable))

sem_out <- as.data.frame(c(sem_out, sem_out_mean))
rownames(sem_out) <- Pops
colnames(sem_out) <- paste("SE", name, sep = "_")

table_out <- cbind(out, sem_out)

return(table_out)
}




###### basic statistics reported from poppr and their SEs
poppr2hierfstat_out <- function(obj, variable){
  
obj_list <- seppop(obj)
  
stats_poppr <- list()
for(i in 1: length(obj_list)){
  stats_poppr[[i]] <- locus_table(obj_list[[i]], information = F)
}

table_out <- list()
for(i in 1:length(obj_list))
table_out[[i]] <- stats_poppr[[i]][-nrow(stats_poppr[[1]]), variable]

table_out <- as.matrix(as.data.frame(table_out))
colnames(table_out) <- popNames(obj)

return(table_out)
}

```


### SSR summary
```{r}
library(poppr)
library(hierfstat)
library(reshape2)

## N
N_by_locus <- basic.stats(obj_SSR)[["n.ind.samp"]]
obj_list <- seppop(obj_SSR)
N <- list()
for(i in 1:length(obj_list)){
        N[[i]] <- length(obj_list[[i]]@pop)
}
N <- melt(N)
N_ssr <- c(N[,1], sum(N[,1]))

## na
na_by_locus <- poppr2hierfstat_out(obj_SSR, "allele")
na_ssr <- table_out(obj_SSR, na_by_locus, "na")


## uHe
uHe_by_locus <- poppr2hierfstat_out(obj_SSR, "Hexp")
uHe_ssr <- table_out(obj_SSR, uHe_by_locus, "uHe")

## Ho
Ho_by_locus <- basic.stats(obj_SSR)[["Ho"]]
Ho_ssr <- table_out(obj_SSR, Ho_by_locus, "Ho")

## ne
ne_by_locus_Hs <- 1 / (1 - (basic.stats(obj_SSR)[["Hs"]]))
ne_Hs_ssr <- table_out(obj_SSR, ne_by_locus_Hs, "ne")

## ## ne
## ne_by_locus_He <- 1 / (1 - (basic.stats(obj_SSR)[["Hs"]]))
## ne_Hs <- table_out(obj_SSR, ne_by_locus, "ne")

## Fis
Fis_by_locus <- basic.stats(obj_SSR)[["Fis"]]
Fis_ssr <- table_out(obj_SSR, Fis_by_locus, "Fis") ## better use boot.ppfis

## Private alleles
PA <- rowSums(poppr::private_alleles(obj_SSR, count.alleles=F))
PA <- data.frame(PA)
PA_ssr <- rbind(PA, Total=colSums(PA))
```

```{r include=T}
uHe_by_locus_tidy <- gather(as.data.frame(uHe_by_locus), 
                            key = "population",
                            value = "uHe")

uHe_by_locus_tidy %>% 
ggplot(aes(x = population, y = uHe) ) +
  geom_violin() +
  # geom_boxplot(width=.1) +
  stat_summary(fun.y=mean, 
                 geom="point") +
  ggtitle("uHe distribution and mean value") +
  theme_classic()

```



### SNP summary
```{r}
library(poppr)
library(hierfstat)
library(reshape2)

## N
N_by_locus <- basic.stats(obj_SNP)[["n.ind.samp"]]
obj_list <- seppop(obj_SNP)
N <- list()
for(i in 1:length(obj_list)){
        N[[i]] <- length(obj_list[[i]]@pop)
}
N <- melt(N)
N_snp <- c(N[,1], sum(N[,1]))

## na
na_by_locus <- poppr2hierfstat_out(obj_SNP, "allele")
na_snp <- table_out(obj_SNP, na_by_locus, "na")


## uHe
uHe_by_locus <- poppr2hierfstat_out(obj_SNP, "Hexp")
uHe_snp <- table_out(obj_SNP, uHe_by_locus, "uHe")

## Ho
Ho_by_locus <- basic.stats(obj_SNP)[["Ho"]]
Ho_snp <- table_out(obj_SNP, Ho_by_locus, "Ho")

## ne
ne_by_locus_Hs <- 1 / (1 - (basic.stats(obj_SNP)[["Hs"]]))
ne_Hs_snp <- table_out(obj_SNP, ne_by_locus_Hs, "ne")

## ## ne
## ne_by_locus_He <- 1 / (1 - (basic.stats(obj_SNP)[["Hs"]]))
## ne_Hs <- table_out(obj_SNP, ne_by_locus, "ne")

## Fis
Fis_by_locus <- basic.stats(obj_SNP)[["Fis"]]
Fis_snp <- table_out(obj_SNP, Fis_by_locus, "Fis") ## better use boot.ppfis

## Private alleles

PA_SNP_p <- poppr::private_alleles(obj_SNP, count.alleles=F)
if(is.numeric(PA_SNP_p)){
  PA_SNP <- rowSums(PA_SNP_p)
  PA_SNP <- data.frame(PA)
  PA <- rbind(PA, Total=colSums(PA_SNP))
}else {
  PA_SNP <- rep(0, length(popNames(obj_SNP)) + 1 )
}

```

## uHe
```{r uHe}
uHe_ssr$Cohort <- rownames(uHe_ssr)
uHe_snp$Cohort <- rownames(uHe_snp)

uHe_ssr$marker <- "11 SSR loci"
uHe_snp$marker <- "140 SNP loci"

#bind dfs
uHe_df <- rbind(uHe_ssr[-3,], uHe_snp[-3,])

uHe_p <- uHe_df %>% 
      ggplot(aes(x=Cohort, y=uHe, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_wrap(~ marker) +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) 

uHe_p + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11))+
  ylab(expression(H[E]))

# SHOW WHAT IS THE SAMPLE SIZE!!!
```

## Ho
```{r Ho}
Ho_ssr$Cohort <- rownames(Ho_ssr)
Ho_snp$Cohort <- rownames(Ho_snp)

Ho_ssr$marker <- "11 SSR loci"
Ho_snp$marker <- "140 SNP loci"

#bind dfs
Ho_df <- rbind(Ho_ssr[-3,], Ho_snp[-3,])

Ho_p <- Ho_df %>% 
      ggplot(aes(x=Cohort, y=Ho, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_wrap(~ marker) +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) 

Ho_p + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11)) +
  ylab(expression(H[O]))

# SHOW WHAT IS THE SAMPLE SIZE!!!
```

## Fis
```{r Fis}
Fis_ssr$Cohort <- rownames(Fis_ssr)
Fis_snp$Cohort <- rownames(Fis_snp)

Fis_ssr$marker <- "11 SSR loci"
Fis_snp$marker <- "140 SNP loci"

#bind dfs
Fis_df <- rbind(Fis_ssr[-3,], Fis_snp[-3,])

Fis_p <- Fis_df %>% 
      ggplot(aes(x=Cohort, y=Fis, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_wrap(~ marker) +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) 

Fis_p + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11)) +
  ylab(expression(F[IS]))

# SFisW WHAT IS THE SAMPLE SIZE!!!
```

## Fst - Gst
```{r}
library(hierfstat)
fst_ssr <- genet.dist(obj_SSR, method = "Nei87")
fst_snp <- genet.dist(obj_SNP, method = "Nei87")
```

## Gst Hedrick

## Nei D

## AMOVA

## LD
```{r eval=F}
library(genetics)

obj_SNP_loci <- genind2loci(obj_SNP)

LD_SNP <- makeGenotypes(obj_SNP_loci[,-1]) %>% 
          genetics::LD()
```

```{r eval=F}
LD_df <- LD_SNP$`P-value`

ind <- which(upper.tri(LD_df, diag = TRUE), 
             arr.ind = TRUE)

nn <- dimnames(LD_df)

LD_df_long <- na.omit(
  data.frame(locus1= nn[[1]] [ ind[,1]],
             locus2= nn[[2]] [ ind[,2]],
             pvalue= LD_df [ ind])
)

# significant LD after multiple test correction
LD_df_long$p.adjust <- p.adjust(LD_df_long$pvalue,
                                method="holm")

LD_sig <- filter(LD_df_long, p.adjust < 0.05)
```

## export to fstat --> NeEstimator
```{r}
# Both data sets have issues with physical LD
# First subset them, then use NeEstimator

## subset SSR data set
all_loci <- locNames(obj_SSR) #vector of all loci
removeloc <- "SF1" #loci to remove
keeploc <- setdiff(all_loci, removeloc) #loci to keep
obj_SSR_ne <- obj_SSR[loc = keeploc]

obj_SSR_hier <- genind2hierfstat(obj_SSR_ne)

write.fstat(dat = obj_SSR_hier, 
            fname = "../data/LGM_GR_Abies_SSR.dat")

## subset SNP data set based on previous analysis
to.remove <- c("contig02161-64", "contig18676-31", "contig21141-311", "newcontig05502-128", "newcontig07067-521", "newcontig08206-176", "newcontig30715-525")
loci_int <- names(nAll(obj_SNP))
obj_SNP_no_LD <- obj_SNP[,loc = setdiff(loci_int, to.remove)]

obj_SNP_hier <- genind2hierfstat(obj_SNP_no_LD)

write.fstat(dat = obj_SNP_hier, 
            fname = "../data/LGM_GR_Abies_SNP.dat")
```

```{r}
# alternative export to neestimator for SSR data set
all_loci <- locNames(obj_SSR) #vector of all loci
removeloc <- "Aat01" #loci to remove
keeploc <- setdiff(all_loci, removeloc) #loci to keep
obj_SSR_ne2 <- obj_SSR[loc = keeploc]

obj_SSR_hier2 <- genind2hierfstat(obj_SSR_ne2)

write.fstat(dat = obj_SSR_hier2, 
            fname = "../data/LGM_GR_Abies_SSR_2.dat")
```

## Ne

## Comparisons with Krajmerova
```{r}
df_kraj <- data.frame(
        ar = 4.29,
        par = 0.03,
        ne = 2.69,
        he = 0.562
)

df_gr <- data.frame(
        ar = c(5.9, 5.6),
        par = c(0.21, 0.18),
        ne = c(4.39, 4.42),
        he = c(0.589, 0.593)
)

df_comp <- rbind(df_kraj, df_gr)
df_comp$study <- c("Krajmerová et al. \n 2015 (7 SSRs)", "Current study \n (11 SSRs)", "Current study \n (11 SSRs)")
df_comp$Cohort <- c("GR_A", "GR_A", "GR_NR")
```

```{r}
ar_comp_p <- df_comp %>% 
      ggplot(aes(x=Cohort, y=ar, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_grid(~ study, scales="free_x",
                       space = "free_x") +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) 

ar_comp_p + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11)) +
        ylab("AR") +
    ggtitle("Allelic Richness (g=44) \nMt Olympus region")
```

```{r}
par_comp_p <- df_comp %>% 
      ggplot(aes(x=Cohort, y=par, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_grid(~ study, scales="free_x",
                       space = "free_x") +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) 

par_comp_p + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11)) +
        ylab("pAR") +
    ggtitle("Private Allelic Richness (g=44) \nMt Olympus region")
```

```{r}
ne_comp_p <- df_comp %>% 
      ggplot(aes(x=Cohort, y=ne, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_grid(~ study, scales="free_x",
                       space = "free_x") +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) 

ne_comp_p + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11)) +
        ylab(expression(n[e])) +
    ggtitle("Effective Number of Alleles \nMt Olympus region")
```

```{r}
he_comp_p <- df_comp %>% 
      ggplot(aes(x=Cohort, y=he, fill=Cohort)) +
            geom_bar(stat="identity", width = 0.5,
                     color="black") +
            facet_grid(~ study, scales="free_x",
                       space = "free_x") +
      theme_classic() +
      theme(legend.position = "none") +
      scale_fill_manual(values = c("white", "lightgrey")) 

he_comp_p + theme(axis.title = element_text(size = 14), 
    axis.text = element_text(size = 11), 
    axis.text.x = element_text(size = 11), 
    axis.text.y = element_text(size = 13),
    strip.text.x = element_text(size = 11)) +
        ylab(expression(H[E])) +
    ggtitle("Expected Heterozygosity \nMt Olympus region")
```

